---
title: "Assignment1"
output: html_document
---

**1.What is mean total number of steps taken per day?**

First,calculate the total number of steps taken per day:
```{r}
data1<-data.frame("steps"=numeric(61),"date"=character(61),stringsAsFactors=FALSE)
for (i in 1:61){
    data1$steps[i]<-sum((data[(288*(i-1)):(288*i),])[,1],na.rm=TRUE)
    data1$date[i]<-(levels(data$date))[[i]]
}
data1
```

Then,make a histogram of the total number of steps taken each day:
```{r}
hist(data1$steps,main="Histogram of The Total Number of Steps Taken Each Day",
     xlab="Total number of steps taken each day",labels=TRUE,ylim=c(0,35))
```

Last,calculate and report the mean and median of the total number of steps taken per day:
```{r}
summary(data1$steps)
```


**2.What is the average daily activity pattern?**

First,make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
data2<-aggregate(steps~interval,data,mean)
plot(data2$interval,data2$steps,type="l",
     main="Average Daily Activity Pattern",
     ylab="Average number of steps taken",xlab="Interval")
```

Then,in order to answer the question "Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?",run the code:
```{r,results="hide"}
library(dplyr)
```
```{r}
head(arrange(data2,desc(steps)))
```
Result showed that the 5-minute interval of "835" contains the maximum number of steps on average across all the days.


**3.Imputing missing values**

First,calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
nrow(filter(data,is.na(steps)))
```

Thus,there are 2304 missing values in the dataset.

Next,use the mean for that 5-minute interval,create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r,result="hide"}
library(plyr)
```
```{r}
impute.mean<-function(x) replace(x,is.na(x),mean(x,na.rm=TRUE))
data3<-ddply(data,~interval,transform,steps=impute.mean(steps))
data3<-data3[order(data3$date),]
```

Last,make a histogram of the total number of steps taken each day:
```{r}
data3_daily<-aggregate(steps~date,data3,sum)
hist(data3_daily$steps,main="Histogram of the total number of steps taken each day
     \n (with missing data filled in)",xlab="Total number of steps taken each day"
     ,label=TRUE,ylim=c(0,40))
```

And calculate and report the mean and median total number of steps taken per day:
```{r}
summary(data3_daily$steps)
```

The mean and the median total number of steps taken per day are both 10770.

Then,compare this histogram with that from the first part of the assignment:
```{r}
par(mfrow=c(1,2),mar=c(5,5,5,4))
hist(data1$steps,main="Histogram of the total number \n of steps taken each day",
     xlab="Total number of steps taken each day",labels=TRUE,ylim=c(0,40))
hist(data3_daily$steps,
     main="Histogram of the total number \n of steps taken each day \n (with missing data filled in)",
     xlab="Total number of steps taken each day",
     label=TRUE,ylim=c(0,40))
```

Plots show that after imputing missing data,the frequency of total number of steps taken each day between 0 and 5000 decreases from 13 to 5, and the frequency of total number of steps taken each day between 10000 and 15000 increases from 28 to 36. 


**4.Are there differences in activity patterns between weekdays and weekends?**

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data3["day_of_week"]<-NA
for (i in 1:nrow(data3)){
    if (weekdays(as.Date(data3$date[i]))=="Saturday" 
        |weekdays(as.Date(data3$date[i]))=="Sunday")
        data3$day_of_week[i]<-"weekend" else
        data3$day_of_week[i]<-"weekday"
}
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r,results="hide"}
library(lattice)
```
```{r}
data4<-aggregate(steps~day_of_week+interval,data3,mean)
xyplot(steps~interval|day_of_week,data=data4,layout=c(1,2),type="l")
```